---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.7.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

<!-- #region nbgrader={"grade": false, "grade_id": "cell-0a3a16e4560c3222", "locked": true, "schema_version": 3, "solution": false, "task": false} -->

# STAT 201 Project: Vancouver Crime

<!-- #endregion -->

## Introduction

### Possible research question

*Is the proportion of bike theft crime compared to other crime different in two different neighbourhoods between 2003 and 2010 and had this changed from 2011 to 2018?*

- **Random Variable:** Bike theft as a proportion of total crime

- **Category 1: Neighbourhoods** (Central Business District vs Kitsilano)

- **Category 2:** Time periods (2003-2010 vs 2011-2018)

### Dataset: 
The dataset we will be using is the Vancouver crime dataset provided by: <https://geodash.vpd.ca/opendata/> . It contains the years, classifications, times, and locations of various crimes that have occurred in Vancouver starting from 2003 up to present.

### Background: 
Crime rates in Vancouver are constantly fluctuating. The COVID-19 pandemic in particular has drastically affected crime rates in the city of Vancouver. Local police recorded a 21.7% rise in serious assaults in the first half of 2020. Since most businesses closed and were left unattended, break-ins to commercial buildings increased by 47.9%. As restrictions loosened in late 2020, breaking and entering crimes began to subside, as well as domestic assault.

### Initial Code
#### Reading Data

```{r message=FALSE,warning=FALSE}
set.seed(8888)
library(readr)
library(tidyverse)
library(repr)
library(digest)
library(todor)
# library(infer)
# library(gridExtra)
# library(cowplot)

myurl <- "https://raw.githubusercontent.com/NicTavares/VancouverCrimeStats/master/crimedata_csv_all_years.csv"

crime_data<-read_csv(url(myurl))
head(crime_data)
```

#### Cleaning the Data

```{r}
# Selecting desired variables
crime_data_clean = subset(crime_data, select = c(TYPE, YEAR, NEIGHBOURHOOD))

# Removing NA
crime_data_clean = crime_data_clean %>% filter(!is.na(TYPE)) %>% filter(!is.na(YEAR)) %>% filter(!is.na(NEIGHBOURHOOD))

# Filtering for the chosen neighbourhoods
crime_data_clean = crime_data_clean %>% filter(NEIGHBOURHOOD == 'Central Business District' | NEIGHBOURHOOD == 'Kitsilano')
head(crime_data_clean)
```

#### Basic Plots

```{r}
# Summarising the incidence data for TYPE and NEIGHBOURHOOD
crime_count <- crime_data_clean %>% 
  group_by(TYPE) %>% 
  summarize(Count = n())

neighbourhood_count <- crime_data_clean %>% 
  group_by(NEIGHBOURHOOD) %>% 
  summarize(Count = n())

# Plotting total crimes across years
year_plot <- crime_data_clean %>% 
  ggplot() + 
  geom_histogram(aes(x = YEAR), bins = 18) + 
  labs(title = "Total crime across years", x = "Year", y = "Number of Crimes")

# Plotting the incidence of bike theft across years
Bike_Plot <- crime_data_clean %>% 
  filter(TYPE == "Theft of Bicycle") %>%
  group_by(NEIGHBOURHOOD) %>% 
  summarize(count = n()) %>%
  ggplot() + 
  geom_col(aes(x = NEIGHBOURHOOD, y = count,fill=NEIGHBOURHOOD)) + 
  labs(title = "Incidence of Bike Theft", x = "Neighbourhood", y = "Incidence") + 
  theme(axis.text.x = element_text(angle = 90),legend.position = "none")

# Plotting the incidence of each crime type across the total time
crime_count_plot <- crime_count %>% 
  ggplot() + 
  geom_col(aes(x = TYPE, y = Count)) + 
  labs(title = "Incidence of Different Crime Types", 
       x = "Type of Crime", y = "Incidence",
       subtitle = "Kitsilano and Central Business District") + 
  theme(axis.text.x = element_text(angle = 90))

# Plotting the incidence of all crime types per neighbourhood chosen across the total time
neighbourhood_count_plot <- neighbourhood_count %>% 
  ggplot() + 
  geom_col(aes(x = NEIGHBOURHOOD, y = Count, fill = NEIGHBOURHOOD)) + 
  labs(title = "Incidence of All Crime Types per Neighbourhood", 
       x = "Neighbourhood", y = "Incidence") + 
  theme(axis.text.x = element_text(angle = 90),legend.position = "none")


print(crime_count_plot)
print(year_plot)
print(neighbourhood_count_plot)
print(Bike_Plot)
```

#### Sampling

```{r}
# Calculating all crime incidents across all years
n = crime_data_clean %>% summarize(n = n())
n

# Sampling from the total crime data
sample = crime_data_clean %>% 
  infer::rep_sample_n(size = 25000) %>% 
  ungroup() %>% 
  select(-replicate)

head(sample)

# Creating the subsets of the sample of interest
# Broken down by neighbourhood and time period
sample_2003_2010_CBD = sample %>% 
  filter(YEAR >= 2003 & YEAR <= 2010 & NEIGHBOURHOOD == 'Central Business District')

sample_2011_2018_CBD = sample %>% 
  filter(YEAR >= 2011 & YEAR <= 2018 & NEIGHBOURHOOD == 'Central Business District')

sample_2003_2010_kits = sample %>%
  filter(YEAR >= 2003 & YEAR <= 2010 & NEIGHBOURHOOD == 'Kitsilano')

sample_2011_2018_kits = sample %>% 
  filter(YEAR >= 2011 & YEAR <= 2018 & NEIGHBOURHOOD == 'Kitsilano')

head(sample_2011_2018_kits)
```

#### Initial Proportion Calculation

```{r}
# Calculating the total ocurrence of crimes per each sample subset
kits_2003_2010_n = sample_2003_2010_kits %>%
  group_by(TYPE) %>% 
  summarize(n = n())
kits_2011_2018_n = sample_2011_2018_kits %>%
  group_by(TYPE) %>%
  summarize(n = n())
CBD_2003_2010_n = sample_2003_2010_CBD %>%
  group_by(TYPE) %>% 
  summarize(n = n())
CBD_2011_2018_n = sample_2011_2018_CBD %>%
  group_by(TYPE) %>% 
  summarize(n = n())

#The proportion of total crimes in Kitsilano from 2003-2010 that were Bicycle theft
total_crimes_kits_2003 = sum(kits_2003_2010_n$n)
bike_theft_kits_2003 = kits_2003_2010_n %>% 
  filter(TYPE == 'Theft of Bicycle')
kits_2003_2010_prop = bike_theft_kits_2003$n/total_crimes_kits_2003

#Proportion for kits 2011 to 2018
total_crimes_kits_2011 = sum(kits_2011_2018_n$n)
bike_theft_kits_2011 = kits_2011_2018_n %>%
  filter(TYPE == 'Theft of Bicycle')
kits_2010_2018_prop = bike_theft_kits_2011$n/total_crimes_kits_2011

#Proportion for CBD 2003 to 2010
total_crimes_CBD_2003 = sum(CBD_2003_2010_n$n)
bikes_theft_CBD_2003 = CBD_2003_2010_n %>%
  filter(TYPE == 'Theft of Bicycle')
CBD_prop_2003 = bikes_theft_CBD_2003$n/total_crimes_CBD_2003

#Proportion for CBD 2011 to 2018
total_crimes_CBD_2011 = sum(CBD_2003_2010_n$n)
bikes_theft_CBD_2011 = CBD_2011_2018_n %>%
  filter(TYPE == 'Theft of Bicycle')
CBD_prop_2011 = bikes_theft_CBD_2011$n/total_crimes_CBD_2011

# Outputting results
cat("Proportion of crimes in Kitsilano that were Bike Theft from 2003-2010:",kits_2003_2010_prop,"\nProportion of crimes in Kitsilano that were Bike Theft from 2011-2018:", kits_2010_2018_prop,"\nProportion of crimes in the Central Business District that were Bike Theft from 2003-2010:",CBD_prop_2003,"\nProportion of crimes in the Central Business District that were Bike Theft from 2011-2018:",CBD_prop_2011)
```

## Methods
### In what ways is this report trustworthy?

*Continue by explaining why the plot(s) and estimates that you produced are not enough to give to a stakeholder, and what you should provide in addition to address this gap. Make sure your plans include at least one hypothesis test and one confidence interval, with at least one of these being calculated using both bootstrapping and asymptotics for comparison.*

The initial estimates here are not trustworthy because they are based only on point estimates from our sample. It is possible that while our point estimates are different they are not different enough to believe that the actual population parameters are truly different. We would still need to perform hypothesis testing to determine our confidence level that the true proportions are different.

### What do you expect to find?

We expect to find that certain crimes such as assault may have increased over time while others may have decreased. We also expect the total number of crimes to increase over the years. ← mention neighbourhood predictions 

**for the bike crime question I would expect bike crimes as a proportion of total crime to be different across different neighbourhoods -Nic**

### What impact could such findings have?

The results of this project could help inform the police on how to allocate their resources the most effectively (i.e., have higher staff in X neighbourhood). Our results may indicate which crimes are becoming more frequent in specified neighbourhoods, information that the Vancouver Police Department could use to strategically schedule staff based on their abilities and experiences.

**Impact for the bike crime as a proportion of crime could something similar to this - it could inform police resource allocation and city budget allocation. For example installing more cameras and bike racks in certain locations or installing bike cage rentals in certain locations -Nic**

### What future questions could this lead to?

A future question this project could inform is "Why has the occurrence of crime x increased/decreased over time?"

**Similarily it could inform the question of why is bike crime greater/lesser/same in this area versus other areas? -Nic**

## References

(1) The Canadian Press. (2020, September 17). Vancouver police record increase in violent crime, decrease in property crime in the first half of 2020. CBCnews. <https://www.cbc.ca/news/canada/british-columbia/vancouver-police-increase-violent-crime-decrease-property-crime-1.5728664>.

(2) Uguen-Csenge, E. (2021, June 27). Here's what happened to crime rates in Vancouver during the pandemic \| CBC News. CBCnews. <https://www.cbc.ca/news/canada/british-columbia/here-s-what-happened-to-crime-rates-in-vancouver-during-the-pandemic-1.6078015>.
